### 2023-03-14 manuscript.tex
Submission (#3) [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/c32563406547c62b8e0e04e565d02aabdeac9843.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/c32563406547c62b8e0e04e565d02aabdeac9843/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/4418809035

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/af028ce9536fb8f3682322c0e67b0e9e211b5ff8/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/af028ce9536fb8f3682322c0e67b0e9e211b5ff8}{MMV-Lab/im2im-paper@af028ce}
on March 14, 2023.

@@

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. Deep learning based methods have achieved huge success in 2D semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{29},\protect\hyperlink{ref-XAffSYIR}{30}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{31}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different hold-out test sets, the model achieved F1-score, 0.883 and 0.888 on test set A and test set B, respectively. The performance was competitive comparing to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{30}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

@@

\hypertarget{d-semantic-segmentation-of-organelles-from-electron-microscopy-images}{%
\subsubsection{3D semantic segmentation of organelles from electron microscopy images}\label{d-semantic-segmentation-of-organelles-from-electron-microscopy-images}}

Semantic segmentation in 3D biomedical image analysis application is not a simple generalization from 2D models by switching 2D operations with 3D operations, but with many practical challenges. Large GPU footprint is one of the biggest issues, which makes many training strategies common in 2D not feasible in 3D, e.g.~limited mini-batch size. \emph{MMV\_Im2Im} is able to take advantage of state-of-the-art ML engineering methods to efficiently handle 3D problems. For example, by using effective half-precision training, one can greatly reduce GPU memory workload for each sample and therefore increase the batch size. When multiple GPUs are available, it is also possible to easily take advantage of the additional resources to scale up the training to multiple GPU cards, even multiple GPU nodes. Here, we trained a 3D model to segment seven different types of organelles (e.g., cell, mitochondrion, alpha granule, etc.) from SBF-SEM image volumes {[}\protect\hyperlink{ref-3L5MIyIB}{32}{]}. Example results can be found in Figure \ref{fig:3dseg}. The prediction still suffered from considerable errors, very likely due to the limited training data (only one image).

\begin{figure}
\hypertarget{fig:3dseg}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/semantci_seg3d.png}
\caption{Result of 3D semantic segmentation results on z-slice 30, 60, 90. From left to right: raw images, ground truth, predictions.}\label{fig:3dseg}
}
\end{figure}

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports \emph{EmbedSeg}-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the objects, compared to other FCN-type models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{33},\protect\hyperlink{ref-14h90Vfg0}{34}{]} (which struggles with handling different shapes, such as elongated concave objects), and CellPose {[}\protect\hyperlink{ref-TugPkOLy}{35}{]} or SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{36}{]} (which are not straightforward to generalize to higher dimensions). For example, different from the others, \emph{EmbedSeg}-type models are even able to generate instance segmentation where each instance contains multiple connected components. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{37}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvment. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone for \emph{EmbedSeg}. For 3D anisotropic microscopy image, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{38}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even not applicable. In this scenario, it is straighforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches and pre-calculated object centers around each individual instance. This may generate a massive amount of patches on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an Embedseg model training only with patches centered around an instance, the model may suffer from degraded performance during inference when there are a large amound of background pixels without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Finally, our improved \emph{Embedseg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance have to be annotated. The exclusion mask can effectively avoid this bottleneck. Another extension comparing to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference in practice. As a demonstration, we applied \emph{EmbedSeg}-like models to a 2D problem of segmenting \emph{C. elegans} from widefield images {[}\protect\hyperlink{ref-138foKNOh}{39}{]}, as well as a 3D problem of nuclear segmentation from fluorescent and brightfield images from the hiPS single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}.

@@

For a 3D demonstration, we conducted three different experiments, generating the nuclear segmentation from the dataset as input, and the real H2B images as the training target.

In the 2D case, we extracted the middle Z-slice from NPM1 images as the training target, while using the NPM1 segmentation results as the input binary mask. With the paired ``mask + microscopy image'' data, we could train the model in a supervised fashion, or randomly shuffle the data to simulate the situation without paired data and train the model in an unsupervised fashion. Example results can be found in Figure \ref{fig:synthetic}.

\begin{figure}
\hypertarget{fig:synthetic}{%
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{images/synthetic.png}
\caption{Example results of (A) 2D synthetic fluorescent images of NPM1 and (B) 3D synthetic fluorescent images of H2B (middle z-slices of two example z-stacks).}\label{fig:synthetic}
}
\end{figure}

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low quality and high quality images are required for supervision, even though such pair data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}. Other methods, such as unsupervised learning based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{41}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.

In this example, we presented an image denoising demonstration with sample data from {[}\protect\hyperlink{ref-12G712Zky}{42}{]}. The goal was to increase the quality of low signal-to-noise ratio (SNR) images of nucleus-stained flatworm (Schmidtea mediterranea) and lightsheet images of Tribolium castaneum (red flour beetle) embryos. The models were trained with paired data acquired with low and high laser intensity on fixed samples, and then applied on live imaging data. For the nucleus-stained flatworm data (a test set of 20 images are available), the model achieved pearson correlation of 0.923 ± 0.029 and structural similarity of 0.627 ± 0.175. Based on the results in Figure \ref{fig:denoising}, it can be observed that the low SNR images can be greatly improved. Systematic quantitative evaluations would be necessary to confirm the biological validity, but beyond the scope of this paper.

@@

In general, there are mainly four major deep learning frameworks that could be applied to microscopy image-to-image transformation: supervised learning with a fully convolutional networks (FCN) type models, supervised learning with pix2pix type models, unsupervised learning to learn mapping between visual domains, and Self2Self-type self-supervised learning {[}\protect\hyperlink{ref-srrRoFXR}{43}{]}. The major difference between FCN based supervised learning and pix2pix based supervised learning is that the pix2pix framework extends an FCN model with an adversarial head as a discriminator to further improves the realism of the prediction. The major difference between the unsupervised framework and the self-supervised framework is that the unsupervised methods still requires examples of the target images, even though the source images and target images do not need to be from the same sample or pixel-wise aligned. But, the self-supervised framework would only need the original images, which could be really helpful when it is impossible to acquire the target images (e.g., there is no truely noise-free or artifact-free image).

Currently, both the FCN-type and pix2pix-type supervised frameworks are well supported in the \emph{MMV\_Im2Im} package. Since our package is designed in a very generic way, it is possible to continuously expand the functionalities when available (ideally with community contributions). For example, diffusion models {[}\protect\hyperlink{ref-1A3yurr7m}{44}{]} can be thought of a modern extension of the pix2pix-type framework and therefore are within our horizon to include into \emph{MMV\_Im2Im}. For the unsupervised framework, only CycleGAN-type methods are supported. We are planning to extend the unsupervised framework with Imaginaire (\url{https://github.com/NVlabs/imaginaire}), which will greatly extend the applicability of \emph{MMV\_Im2Im} (e.g., learning the transformation from one single image to another single image or one set of images to another set of images). Meanwhile, supporting the self-supervised framework will be our next major milestone.

@@

One of main directions for extending \emph{MMV\_Im2Im} is to pre-pack common bioimaging datasets as a \emph{Dataset} module, so that users can easily test and benchmarking their method development. Besides the continuous improvement on the functionalities of the package, we also plan to develop two auxillary packages \emph{MMV\_Im2Im\_Auto} and \emph{MMV\_Im2Im\_Active}. In specific, when you have a reasonable amount of training data, MMV\_Im2Im\_Auto will take advantage of fact that \emph{MMV\_Im2Im} is fully configurable with yaml files, and automatically generate a set of potentially good configurations, then find the optimal solution for you. On the other hand, when you only have very limited training data, or even with only pseudo ground truth, \emph{MMV\_Im2Im\_Active} will help to build preliminary models from the limited training data, and gradually refine the model with human-in-the-loop by active learning. All the packages will also be wrapped into Napari plugins {[}\protect\hyperlink{ref-YEMgt2T4}{45}{]} to allow no-code operation and therefore be more friendly to users without experience in programming.

@@

\leavevmode\vadjust pre{\hypertarget{ref-45Sirz1X}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{\textbf{A Stochastic Polygons Model for Glandular Structures in Colon Histology Images}
\CSLBlock{Korsuk Sirinukunwattana, David RJ Snead, Nasir M Rajpoot} \emph{IEEE Transactions on Medical Imaging} (2015-11) \url{https://doi.org/gqrg95}
\CSLBlock{DOI: \href{https://doi.org/10.1109/tmi.2015.2433900}{10.1109/tmi.2015.2433900} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/25993703}{25993703}}}

\leavevmode\vadjust pre{\hypertarget{ref-XAffSYIR}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{\textbf{Gland segmentation in colon histology images: The glas challenge contest}
\CSLBlock{Korsuk Sirinukunwattana, Josien PW Pluim, Hao Chen, Xiaojuan Qi, Pheng-Ann Heng, Yun Bo Guo, Li Yang Wang, Bogdan J Matuszewski, Elia Bruni, Urko Sanchez, \ldots{} Nasir M Rajpoot} \emph{Medical Image Analysis} (2017-01) \url{https://doi.org/c5t7}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.media.2016.08.008}{10.1016/j.media.2016.08.008} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/27614792}{27614792}}}

\leavevmode\vadjust pre{\hypertarget{ref-tQhnZyjK}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{\textbf{A method for normalizing histology slides for quantitative analysis}
\CSLBlock{Marc Macenko, Marc Niethammer, JS Marron, David Borland, John T Woosley, Xiaojun Guan, Charles Schmitt, Nancy E Thomas} \emph{2009 IEEE International Symposium on Biomedical Imaging: From Nano to Macro} (2009-06) \url{https://doi.org/bmbj4h}
\CSLBlock{DOI: \href{https://doi.org/10.1109/isbi.2009.5193250}{10.1109/isbi.2009.5193250}}}

\leavevmode\vadjust pre{\hypertarget{ref-3L5MIyIB}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{\textbf{Dense cellular segmentation for EM using 2D--3D neural network ensembles}
\CSLBlock{Matthew D Guay, Zeyad AS Emam, Adam B Anderson, Maria A Aronova, Irina D Pokrovskaya, Brian Storrie, Richard D Leapman} \emph{Scientific Reports} (2021-01-28) \url{https://doi.org/gqdxxg}
\CSLBlock{DOI: \href{https://doi.org/10.1038/s41598-021-81590-0}{10.1038/s41598-021-81590-0} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/33510185}{33510185} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7844272}{PMC7844272}}}

\leavevmode\vadjust pre{\hypertarget{ref-tIIG2f8K}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{\textbf{Cell Detection with Star-Convex Polygons}
\CSLBlock{Uwe Schmidt, Martin Weigert, Coleman Broaddus, Gene Myers} \emph{Medical Image Computing and Computer Assisted Intervention -- MICCAI 2018} (2018) \url{https://doi.org/ggnzqb}
\CSLBlock{DOI: \href{https://doi.org/10.1007/978-3-030-00934-2_30}{10.1007/978-3-030-00934-2\_30}}}

\leavevmode\vadjust pre{\hypertarget{ref-14h90Vfg0}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{\textbf{Star-convex Polyhedra for 3D Object Detection and Segmentation in Microscopy}
\CSLBlock{Martin Weigert, Uwe Schmidt, Robert Haase, Ko Sugawara, Gene Myers} \emph{2020 IEEE Winter Conference on Applications of Computer Vision (WACV)} (2020-03) \url{https://doi.org/gjp4g9}
\CSLBlock{DOI: \href{https://doi.org/10.1109/wacv45572.2020.9093435}{10.1109/wacv45572.2020.9093435}}}

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/c32563406547c62b8e0e04e565d02aabdeac9843/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/c32563406547c62b8e0e04e565d02aabdeac9843}{MMV-Lab/im2im-paper@c325634}
on March 14, 2023.

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports \emph{EmbedSeg}-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the objects, compared to other FCN-type models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{29},\protect\hyperlink{ref-14h90Vfg0}{30}{]} (which struggles with handling different shapes, such as elongated concave objects), and CellPose {[}\protect\hyperlink{ref-TugPkOLy}{31}{]} or SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{32}{]} (which are not straightforward to generalize to higher dimensions). For example, different from the others, \emph{EmbedSeg}-type models are even able to generate instance segmentation where each instance contains multiple connected components. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{33}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvment. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone for \emph{EmbedSeg}. For 3D anisotropic microscopy image, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{34}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even not applicable. In this scenario, it is straighforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches and pre-calculated object centers around each individual instance. This may generate a massive amount of patches on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an Embedseg model training only with patches centered around an instance, the model may suffer from degraded performance during inference when there are a large amound of background pixels without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Finally, our improved \emph{Embedseg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance have to be annotated. The exclusion mask can effectively avoid this bottleneck. Another extension comparing to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference in practice.

In this work, we tested on both 2D and 3D instance segmentation problems. Going from 2D to 3D is not a simple generalization from 2D models by switching 2D operations with 3D operations, but with many practical challenges. Large GPU footprint is one of the biggest issues, which makes many training strategies common in 2D not feasible in 3D, e.g.~limited mini-batch size. \emph{MMV\_Im2Im} is able to take advantage of state-of-the-art ML engineering methods to efficiently handle 3D problems. For example, by using effective half-precision training, one can greatly reduce GPU memory workload for each sample and therefore increase the batch size. When multiple GPUs are available, it is also possible to easily take advantage of the additional resources to scale up the training to multiple GPU cards, even multiple GPU nodes. As a demonstration, we applied \emph{EmbedSeg}-like models to a 2D problem of segmenting \emph{C. elegans} from widefield images {[}\protect\hyperlink{ref-138foKNOh}{35}{]}, as well as a 3D problem of nuclear segmentation from fluorescent and brightfield images from the hiPS single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}.

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. Deep learning based methods have achieved huge success in 2D semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{36},\protect\hyperlink{ref-XAffSYIR}{37}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{38}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different hold-out test sets, the model achieved F1-score, 0.883 and 0.888 on test set A and test set B, respectively. The performance was competitive comparing to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{37}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

@@

\hypertarget{comparing-semantic-segmentation-and-instance-segmentation-of-organelles-from-3d-confocal-microscopy-images}{%
\subsubsection{Comparing semantic segmentation and instance segmentation of organelles from 3D confocal microscopy images}\label{comparing-semantic-segmentation-and-instance-segmentation-of-organelles-from-3d-confocal-microscopy-images}}

We did a special comparison in this section to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{24}{]}. There are multiple channels in each 3D images, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin in this case). The original fibrillarin segmentation released with the dataset is a semantic segmentation (0=background, 1=fibrillarin). With the additional cell segmentation available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the original 3D segmentation segmentation ground truth into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals. Meanwhile, the instance segmentation can group them properly so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to the cell membrane channel or cell segmentation results. This is not a simple grouping step based on distance, since the fibrillarin signals from tightly touching nuclei may exist very close to each other.

\begin{figure}
\hypertarget{fig:3dseg}{%
\centering
\includegraphics[width=0.75\textwidth,height=\textheight]{images/semantic_seg3d.png}
\caption{Comparing 3D semantic segmentation and 3D instance segmentation results on confocal microscopy images of fibrillarin (showing a middle Z-slice of a 3D stack). From left to right: raw fibrillarin image, semantic segmentation, instance segmentation, instance ground truth, a reference images with all the channels (DNA dye in cyan, membrane dye in magenta, and fibrillarin in white) to show all the neighering cells.}\label{fig:3dseg}
}
\end{figure}

@@

@@

For 2D demonstration, we extracted the middle Z-slice from NPM1 images as the training target, while using the NPM1 segmentation results as the input binary mask. With the paired ``mask + microscopy image'' data, we could train the model in a supervised fashion, or randomly shuffle the data to simulate the situation without paired data which can be trained in an unsupervised fashion. Example results can be found in Figure \ref{fig:synthetic}-A. In general, the supervised synthesization can generate more realistic images than the unsupervised model.

For 3D demonstration, we use the 3D H2B images with two different types of input masks. First, we attempted to generate synthetic images from a coarse mask (i.e., only the overall shape of the nucleus, available as nuclear segmentation from the dataset) with both supervised training and unsupervised training. The unsupervised model in \emph{MMV\_Im2Im} uses the CycleGAN-based approaches. So, the unsupervised training is actually already done within the unsupervised segmentation experiments. In other words, the unsupervised model works in a bi-directional way, from real microscopy images to binary mask, and also from binary masks to simulated microscopy images. Here, we just do the inference in a different direction (from binary to simulated microscopy) using the model trained in unsupervised segmentation section. The results are shown in Figure \ref{fig:synthetic}-B (row 1). Similar to the results in 2D demonstration, the unsupervised synthesization can mostly ``paint'' the mask with homogeneous grayscale intensity, while the supervised model can simulate the textures to some extent. We observed for a relatively large mask, it could be chanllenging for a model to fill in sufficient details to simulate real microscopy images (might be improved with diffusion-based models, see Discussions).

We made another attempt with 3D masks containing finer details beyond the overall shapes. So, we employed the H2B structure segmentation results from the dataset (capturing the detailed nuclear components marked by histon H2B) as the input for supervsied synthesization. The result is shown in Figure \ref{fig:synthetic}-B (row 2). Comparing to the synthesization with coarse masks, the images simulated from fine masks exhibit much more realistic appearance. As we can see, it is important to design the solutions with proper data.

\begin{figure}
\hypertarget{fig:synthetic}{%
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{images/synthetic.png}
\caption{Example results of (A) 2D synthetic fluorescent images of nucleoli (via NPM1) and (B) 3D synthetic fluorescent images of H2B (middle z-slices of a z-stack) with a coarse mask and a fine mask as the input.}\label{fig:synthetic}
}
\end{figure}

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low quality and high quality images are required for supervision, even though such pair data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}. Other methods, such as unsupervised learning based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{40}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.

In this example, we presented an image denoising demonstration with sample data from {[}\protect\hyperlink{ref-12G712Zky}{41}{]}. The goal was to increase the quality of low signal-to-noise ratio (SNR) images of nucleus-stained flatworm (Schmidtea mediterranea) and lightsheet images of Tribolium castaneum (red flour beetle) embryos. The models were trained with paired data acquired with low and high laser intensity on fixed samples, and then applied on live imaging data. For the nucleus-stained flatworm data (a test set of 20 images are available), the model achieved pearson correlation of 0.923 ± 0.029 and structural similarity of 0.627 ± 0.175. Based on the results in Figure \ref{fig:denoising}, it can be observed that the low SNR images can be greatly improved. Systematic quantitative evaluations would be necessary to confirm the biological validity, but beyond the scope of this paper.

@@

In general, there are mainly four major deep learning frameworks that could be applied to microscopy image-to-image transformation: supervised learning with a fully convolutional networks (FCN) type models, supervised learning with pix2pix type models, unsupervised learning to learn mapping between visual domains, and Self2Self-type self-supervised learning {[}\protect\hyperlink{ref-srrRoFXR}{42}{]}. The major difference between FCN based supervised learning and pix2pix based supervised learning is that the pix2pix framework extends an FCN model with an adversarial head as a discriminator to further improves the realism of the prediction. The major difference between the unsupervised framework and the self-supervised framework is that the unsupervised methods still requires examples of the target images, even though the source images and target images do not need to be from the same sample or pixel-wise aligned. But, the self-supervised framework would only need the original images, which could be really helpful when it is impossible to acquire the target images (e.g., there is no truely noise-free or artifact-free image).

Currently, both the FCN-type and pix2pix-type supervised frameworks are well supported in the \emph{MMV\_Im2Im} package. Since our package is designed in a very generic way, it is possible to continuously expand the functionalities when available (ideally with community contributions). For example, diffusion models {[}\protect\hyperlink{ref-1A3yurr7m}{43}{]} can be thought of a modern extension of the pix2pix-type framework and therefore are within our horizon to include into \emph{MMV\_Im2Im}. For the unsupervised framework, only CycleGAN-type methods are supported. We are planning to extend the unsupervised framework with Imaginaire (\url{https://github.com/NVlabs/imaginaire}), which will greatly extend the applicability of \emph{MMV\_Im2Im} (e.g., learning the transformation from one single image to another single image or one set of images to another set of images). Meanwhile, supporting the self-supervised framework will be our next major milestone.

@@

One of main directions for extending \emph{MMV\_Im2Im} is to pre-pack common bioimaging datasets as a \emph{Dataset} module, so that users can easily test and benchmarking their method development. We will continue improving the functionalities of the package, such as supporting more models and methods, such as diffusion based models {[}\protect\hyperlink{ref-1A3yurr7m}{43}{]} or Imaginaire (\url{https://github.com/NVlabs/imaginaire}). Besides, we also plan to develop two auxillary packages \emph{MMV\_Im2Im\_Auto} and \emph{MMV\_Im2Im\_Active}. In specific, when you have a reasonable amount of training data, MMV\_Im2Im\_Auto will take advantage of fact that \emph{MMV\_Im2Im} is fully configurable with yaml files, and automatically generate a set of potentially good configurations, then find the optimal solution for you. On the other hand, when you only have very limited training data, or even with only pseudo ground truth, \emph{MMV\_Im2Im\_Active} will help to build preliminary models from the limited training data, and gradually refine the model with human-in-the-loop by active learning. All the packages will also be wrapped into Napari plugins {[}\protect\hyperlink{ref-YEMgt2T4}{44}{]} to allow no-code operation and therefore be more friendly to users without experience in programming.

diffusion based models

@@

\leavevmode\vadjust pre{\hypertarget{ref-tIIG2f8K}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{\textbf{Cell Detection with Star-Convex Polygons}
\CSLBlock{Uwe Schmidt, Martin Weigert, Coleman Broaddus, Gene Myers} \emph{Medical Image Computing and Computer Assisted Intervention -- MICCAI 2018} (2018) \url{https://doi.org/ggnzqb}
\CSLBlock{DOI: \href{https://doi.org/10.1007/978-3-030-00934-2_30}{10.1007/978-3-030-00934-2\_30}}}

\leavevmode\vadjust pre{\hypertarget{ref-14h90Vfg0}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{\textbf{Star-convex Polyhedra for 3D Object Detection and Segmentation in Microscopy}
\CSLBlock{Martin Weigert, Uwe Schmidt, Robert Haase, Ko Sugawara, Gene Myers} \emph{2020 IEEE Winter Conference on Applications of Computer Vision (WACV)} (2020-03) \url{https://doi.org/gjp4g9}
\CSLBlock{DOI: \href{https://doi.org/10.1109/wacv45572.2020.9093435}{10.1109/wacv45572.2020.9093435}}}

\leavevmode\vadjust pre{\hypertarget{ref-TugPkOLy}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{\textbf{Cellpose: a generalist algorithm for cellular segmentation}
\CSLBlock{Carsen Stringer, Tim Wang, Michalis Michaelos, Marius Pachitariu} \emph{Nature Methods} (2020-12-14) \url{https://doi.org/ghrgms}
\CSLBlock{DOI: \href{https://doi.org/10.1038/s41592-020-01018-x}{10.1038/s41592-020-01018-x} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/33318659}{33318659}}}

\leavevmode\vadjust pre{\hypertarget{ref-17Yrl6WGQ}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{\textbf{Splinedist: Automated Cell Segmentation With Spline Curves}
\CSLBlock{Soham Mandal, Virginie Uhlmann} \emph{2021 IEEE 18th International Symposium on Biomedical Imaging (ISBI)} (2021-04-13) \url{https://doi.org/gqrg94}
\CSLBlock{DOI: \href{https://doi.org/10.1109/isbi48211.2021.9433928}{10.1109/isbi48211.2021.9433928}}}

\leavevmode\vadjust pre{\hypertarget{ref-xi8wnibR}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{\textbf{Mask R-CNN}
\CSLBlock{Kaiming He, Georgia Gkioxari, Piotr Dollar, Ross Girshick} \emph{IEEE Transactions on Pattern Analysis and Machine Intelligence} (2020-02-01) \url{https://doi.org/gfxfwn}
\CSLBlock{DOI: \href{https://doi.org/10.1109/tpami.2018.2844175}{10.1109/tpami.2018.2844175} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/29994331}{29994331}}}

\leavevmode\vadjust pre{\hypertarget{ref-XAkgs3Nh}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{\textbf{ERFNet: Efficient Residual Factorized ConvNet for Real-Time Semantic Segmentation}
\CSLBlock{Eduardo Romera, Jose M Alvarez, Luis M Bergasa, Roberto Arroyo} \emph{IEEE Transactions on Intelligent Transportation Systems} (2018-01) \url{https://doi.org/gcs5h7}
\CSLBlock{DOI: \href{https://doi.org/10.1109/tits.2017.2750080}{10.1109/tits.2017.2750080}}}

\leavevmode\vadjust pre{\hypertarget{ref-138foKNOh}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{\textbf{High-Throughput Screen for Novel Antimicrobials using a Whole Animal Infection Model}
\CSLBlock{Terence I Moy, Annie L Conery, Jonah Larkins-Ford, Gang Wu, Ralph Mazitschek, Gabriele Casadei, Kim Lewis, Anne E Carpenter, Frederick M Ausubel} \emph{ACS Chemical Biology} (2009-06-29) \url{https://doi.org/bdwdfc}
\CSLBlock{DOI: \href{https://doi.org/10.1021/cb900084v}{10.1021/cb900084v} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/19572548}{19572548} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2745594}{PMC2745594}}}

\leavevmode\vadjust pre{\hypertarget{ref-45Sirz1X}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{\textbf{A Stochastic Polygons Model for Glandular Structures in Colon Histology Images}
\CSLBlock{Korsuk Sirinukunwattana, David RJ Snead, Nasir M Rajpoot} \emph{IEEE Transactions on Medical Imaging} (2015-11) \url{https://doi.org/gqrg95}
\CSLBlock{DOI: \href{https://doi.org/10.1109/tmi.2015.2433900}{10.1109/tmi.2015.2433900} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/25993703}{25993703}}}
